Data Exploration

### 2019-02-11 src/data_exploration.Rmd
update to one vif + some early maxent setup with dismo (just to test - maxent will be run in a seperate script)

#with gear
vif_gear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
write.csv(vif_gear_cor_apr, "../output/bio/vif/monthly_vifs/vif_gear_cor_apr.csv", row.names = TRUE)

@@

#without gear
vif_nogear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = aprpresab))
write.csv(vif_nogear_cor_apr, "../output/bio/vif/monthly_vifs/vif_nogear_cor_apr.csv", row.names = TRUE)
vif_nogear_cor_apr

@@

#with gear
vif_gear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
write.csv(vif_gear_cor_jun, "../output/bio/vif/monthly_vifs/vif_gear_cor_jun.csv", row.names = TRUE)
vif_gear_cor_jun

#without gear
vif_nogear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
write.csv(vif_nogear_cor_jun, "../output/bio/vif/monthly_vifs/vif_nogear_cor_jun.csv", row.names = TRUE)
vif_nogear_cor_jun

@@

#with gear
vif_gear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
write.csv(vif_gear_cor_oct, "../output/bio/vif/monthly_vifs/vif_gear_cor_oct.csv", row.names = TRUE)
vif_gear_cor_oct

@@

#without gear
vif_nogear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
write.csv(vif_nogear_cor_oct, "../output/bio/vif/monthly_vifs/vif_nogear_cor_oct.csv", row.names = TRUE)
vif_nogear_cor_oct

@@

ok now as per MJ/Chan, count of cell sampled by year, and then cell sampled by month (do this on the data with obs with missing env. data out)

@@

```{r}
library("raster")
library("dismo")
library("rgeos")
```

>>>>

#with gear
vif_gear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
write.csv(vif_gear_cor_apr, "../output/bio/vif/monthly_vifs/vif_gear_cor_apr.csv", row.names = TRUE)

@@

#without gear
vif_nogear_cor_apr <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = aprpresab))
write.csv(vif_nogear_cor_apr, "../output/bio/vif/monthly_vifs/vif_nogear_cor_apr.csv", row.names = TRUE)
vif_nogear_cor_apr

@@

#with gear
vif_gear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
write.csv(vif_gear_cor_jun, "../output/bio/vif/monthly_vifs/vif_gear_cor_jun.csv", row.names = TRUE)
vif_gear_cor_jun

#without gear
vif_nogear_cor_jun <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = junpresab))
write.csv(vif_nogear_cor_jun, "../output/bio/vif/monthly_vifs/vif_nogear_cor_jun.csv", row.names = TRUE)
vif_nogear_cor_jun

@@

#with gear
vif_gear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
write.csv(vif_gear_cor_oct, "../output/bio/vif/monthly_vifs/vif_gear_cor_oct.csv", row.names = TRUE)
vif_gear_cor_oct

@@

#without gear
vif_nogear_cor_oct <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = octpresab))
write.csv(vif_nogear_cor_oct, "../output/bio/vif/monthly_vifs/vif_nogear_cor_oct.csv", row.names = TRUE)
vif_nogear_cor_oct

@@

# chan/mj - samples per cell by year and month. 
ok now as per MJ/Chan, count of cell sampled by year, and then cell sampled by month (do this on the data with obs with missing env. data out)
note this is a summary of the month/year - cells can be sampled more than one, but never more than once at the same month/year

@@

This test will be run on january data

subset the presab dataset to january only

```{r}
janpresabmax <- subset(presab_missingvals, month == "1")
head(janpresabmax)
```

ok so maxent files need to be in a particular format: species name, x (longitude), y (latitude), env variable, env variable,....

Maxent doesn't have a way to deal with 3D locations (it only takes xy coordinates), however it is possible to use dummy xy coordinates (see appendicies of Elith 2011 A statistical explanation of MaxEnt for ecologists). In my case, I will use the 'id' (the unique id assigned to the sample) for x and "cell_id_xyzt" (the unique cell id on xyzt) for y. I will also want to split the dataset by presence and background, so include here

```{r}
janpresabmax <- subset (janpresabmax, select = c(occurrence, originalscientificname, id, cell_id_xyzt, temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_winter))
head(janpresabmax)
```

now split into presence and background

```{r}
janpresmax <- subset(janpresabmax, occurrence == "1")
janpresmax <- subset (janpresmax, select = -c(occurrence))
janabmax <- subset(janpresabmax, occurrence == "0")
janabmax <- subset (janabmax, select = -c(occurrence))
```

ok so for maxent you need to install the jar file. see https://cran.r-project.org/web/packages/dismo/vignettes/sdm.pdf but note that the location of the jar file has changed http://biodiversityinformatics.amnh.org/open_source/maxent/

citation: Steven J. Phillips, Miroslav Dudík, Robert E. Schapire. [Internet] Maxent software for modeling species niches and distributions (Version 3.4.1). Available from url: http://biodiversityinformatics.amnh.org/open_source/maxent/. Accessed on 2019-2-11.

```{r}
system.file("java", package="dismo") #finding location to put maxent jar file
system("java -version")
```

```{r}
library(dismo)
library(rJava)
jar <- paste(system.file(package="dismo"), "C:/Users/bunny/Documents/GitHub/Chapter_1/packrat/lib/x86_64-w64-mingw32/3.5.1/dismo/java/maxent.jar", sep='')
testmax <- maxent(janpresmax, janabmax)
```
